GVN-Hoist: Hoisting Computations from Branches

### 2016-08-09 hoist.tex
fully redundant
reorganize policies and cost model

This paper describes code hoisting, a technique that uses the information
computed for PRE to detect identical computations and has a transformation phase
whose goal differs from PRE: it removes identical computations from different
branches of execution.  These identical computations would not be redundant at
run-time and thus the number of run-time computations is not reduced.  Instead,
the goals of code hoisting are:
\begin{itemize}
\item to reduce the code size of the program;
\item to improve function inlining heuristics: functions become cheaper to
  inline by reducing their code size;
\item to expose more instruction level parallelism: by hoisting identical
  computations to be executed earlier, schedulers can move heavy computations
  earlier in order to avoid pipeline bubbles;
\item to help out-of-order processors with speculative execution of branches: by
  hoisting expressions out of branches, code hoisting can effectively reduce the
  amount of code to be speculatively executed and can reduce the critical path;
\item to reduce register pressure: by moving computations closer to the
  definitions of their operands;
\item to improve passes that do not work well with branches:
  \begin{itemize}
  \item to improve loop vectorization by reducing a loop with control flow to a
    loop with a single BB, should all the instructions in a conditional get
    hoisted and sinked;
  \item to enable more loop invariant code motion (LICM): as LICM does not
    reason about instructions in the context of loops with conditional branches,
    code-hoisting is needed to move instructions out of conditional expressions
    and expose them to LICM.
  \end{itemize}
\end{itemize}

@@

\subsection{Verification}
Performing a set of checks (e.g., consistency of use-defs and SSA semantics) to
establish that program invariants are maintained.

\subsection{Hoisting bottom up with optimistic approach}
\label{subsec:optimistic}
Optimistic means collect the set of all potential redundancies and
then discard the bad ones.

@@

\subsection{Cost models}

@@

\subsection{Other Approaches}

\subsubsection{Optimistic approach}
In this approach, the goal is to maximise the total number of hoistings in the
entire function.  This algorithm is very useful when optimizing for code-size.
We see a lot of computations hoisted (see
Section~\ref{sec:experimental-results}. In order to do that, we collect the GVN
of all the instructions in the function. After that we iterate on the list of
instructions having identical GVN. After that we find the common dominator
dominating all the identical computations and try to hoist the instruction in
that basic block, if feasible. Many times it is not possible to hoist all the
instructions to one common dominator, due to legality/profitability
constraints. In those cases, the algorithmn would partion the list of
computations a partially hoist a set of computations to their common
dominator. For more details see Section-\ref{subsec:partition}.

\subsubsection{pros,cons}
Naturally, the amount of hoisting depends on the generality of the GVN
algorithm. Current llvm GVN implementation finds out equivalence between
dependent instructions to some extent.

>>>>

This paper describes code hoisting, a technique that uses the information
computed for PRE to detect identical computations and has a transformation phase
whose goal differs from PRE: it removes identical computations from different
branches of execution.  These identical computations would not be redundant at
run-time and thus the number of run-time computations is not reduced.  Instead,
the goals of code hoisting are:
\begin{itemize}
\item to reduce the code size of the program;
\item to remove fully redundant instructions: although this is also done by
  GVN-PRE.
\item to improve function inlining heuristics: functions become cheaper to
  inline by reducing their code size;
\item to expose more instruction level parallelism: by hoisting identical
  computations to be executed earlier, schedulers can move heavy computations
  earlier in order to avoid pipeline bubbles;
\item to help out-of-order processors with speculative execution of branches: by
  hoisting expressions out of branches, code hoisting can effectively reduce the
  amount of code to be speculatively executed and can reduce the critical path;
\item to reduce register pressure: by moving computations closer to the
  definitions of their operands;
\item to improve passes that do not work well with branches:
  \begin{itemize}
  \item to improve loop vectorization by reducing a loop with control flow to a
    loop with a single BB, should all the instructions in a conditional get
    hoisted and sinked;
  \item to enable more loop invariant code motion (LICM): as LICM does not
    reason about instructions in the context of loops with conditional branches,
    code-hoisting is needed to move instructions out of conditional expressions
    and expose them to LICM.
  \end{itemize}
\end{itemize}

@@

\subsection{Verifying the post-conditions}
Performing a set of checks (e.g., consistency of use-defs and SSA semantics) to
establish that program invariants are maintained.

\section{Code hoisting policies}
The amount of hoisting depends on the whether we collect information about all
the instructions before finding candidates (optimistic), or on demand
(pessimistic). It also depends on the generality of the GVN algorithm, however,
that analysis is beyond the scope of this paper.

\subsection{Hoisting bottom up with optimistic approach}
\label{subsec:optimistic}
In this approach, the goal is to maximise the total number of hoistings in the
entire function.  This algorithm is very useful when optimizing for code-size.
In order to do that, we collect the GVN of all the instructions in the
function. After that we iterate on the list of instructions having identical
GVNs. After that we find the common dominator dominating all the identical
computations and try to hoist the instruction in that basic block, if
feasible. Many times it is not possible to hoist all the instructions to one
common dominator, due to legality/profitability constraints. In those cases, the
algorithmn would partion the list of computations and partially hoist a set of
computations to their common dominator. For more details see
Section-\ref{subsec:partition}.

@@

\section{Cost models}

@@
